---
title: '**Data Dictionary for Type 1 Hospitals Data Capture Tool**'
author: "B199291"
date: "`r format (Sys.time (), '%d %B, %Y')`"
output:
  pdf_document: default
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```
# **Load packages for dat dictionary creation**
```{r load}
library(dataMeta)
library (tidyverse)
library(here)
```
Read csv funtion to look collected dataset.
```{r data}
CollectedData=read_csv(here("RefData", "CollectedData_final.csv"))
glimpse(CollectedData) 
```
##**Creating Variable Description** 
```{r variable_description}
variable_description <- c("The index column that allows us to link the data 
collected to the original ae_attendances data in the 'RefData' folder.",
"The month that this activity relates to, stored as a date (1st of each month).",
"The Organisation data service (ODS) code for the organisation. If you want to 
know the organisation associated with a particular ODS code, you can look it up 
from the following address: https://odsportal.digital.nhs.uk/Organisation/Search.", 
"The number of attendances for this department type at this organisation for 
this month.","The number of attendances that breached the four-hour target.",  
"The consent from the end-user to process and share the data collected with 
the data capture tool.")
print(variable_description)
```

## **Creating Variable types**
```{r glimpse2}
glimpse(CollectedData) 
```
We have three quantitative values (measured values) variables and three fixed values (allowable values or codes) variables.
```{r variable_type}
variable_type <- c(0, 1, 1, 0, 0,1)
print(variable_type)
```

###**Building Data Linker for connection between Collected Data and Data Dictionary
```{r build_linker}
linker<-build_linker(CollectedData, variable_description, variable_type)
print(linker)
```

## **Data dictionary**
```{r dictionary}
dictionary <- build_dict(my.data = CollectedData, linker = linker)
glimpse(dictionary)
```
```{r notes}
dictionary[5,4]<-"NHS Trust - CAMBRIDGE UNIVERSITY HOSPITALS NHS FOUNDATION TRUST"
dictionary[6,4]<-"NHS Trust - NORFOLK AND NORWICH UNIVERSITY HOSPITALS NHS FOUNDATION TRUST"
dictionary[7,4]<-"NHS Trust - UNIVERSITY HOSPITALS BIRMINGHAM NHS FOUNDATION TRUST "
dictionary[8,4]<-"NHS Trust - EAST CHESHIRE NHS TRUST "
dictionary[9,4]<-"NHS Trust - SALISBURY NHS FOUNDATION TRUST "
dictionary[10,4]<-"NHS Trust - BIRMINGHAM WOMEN'S AND CHILDREN'S NHS FOUNDATION TRUST "
dictionary[11,4]<-"NHS Trust - WIRRAL UNIVERSITY TEACHING HOSPITAL NHS FOUNDATION TRUST "
dictionary[12,4]<-"NHS Trust - LONDON NORTH WEST UNIVERSITY HEALTHCARE NHS TRUST "
dictionary[13,4]<-"NHS Trust - MEDWAY NHS FOUNDATION TRUST "
dictionary[14,4]<-"NHS Trust - WESTON AREA HEALTH NHS TRUST"
dictionary[15,4]<-"NHS Trust - EAST SUFFOLK AND NORTH ESSEX NHS FOUNDATION TRUST "
```
 

### **Let's save the data dictionary for CollectedData to the 'RefData' folder**
```{r w,message = FALSE, warning = FALSE}
glimpse(dictionary)
write_csv(dictionary, here("RefData", "CollectedData_Datadictionary.csv"))
```
## **Append data dictionary to the CollectedData**
### **Create main_string for attributes**
```{r main_string}
main_string <- "This data describes the NHS England accident and emergency 
(A&E) attendances and breaches of four-hour wait time target data of Type 1 
Hospitals from the *NHSRdatasets* package collected by the data capture tool."
main_string
```
### **Incorporate attributes as metada**
```{r complete_dataset}
complete_CollectedData <- incorporate_attr(my.data = CollectedData, 
data.dictionary = dictionary,main_string = main_string)
attributes(complete_CollectedData)$author[1]<-"B199291"
complete_CollectedData
attributes(complete_CollectedData)
```
### **Save the CollectedData with attributes**
```{r save_it}
save_it(complete_CollectedData, here("RefData", "complete_CollectedData"))
```
